---
title: "Lab8"
author: "Valerie Kamgue"
date: "11/24/2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(foreign)
library(fields)
library(classInt)
library(RColorBrewer)
library(sp)
library(spdep)
library(rgdal)
library(ggplot2)
```


##Part1: Median Polish

```{r}
phyto<-read.csv("C:/Users/valer/Desktop/SIE512/Labs/Lab8/Lab82019/phytophthora.csv")
summary(phyto)

```


##Create a plot of the soilm variable and include it in your lab report.

```{r}
ggplot(phyto, aes(x=X, y=Y, size =soilm, col=soilm))+ scale_color_gradient(low="darkgreen", high="yellow")+ geom_point() + ggtitle("Soil Moisture")
```



##Use the medpolish function to perform the median polish on soil moisture content. 
##Include the image in your lab report and describe any spatial patterns in this image

```{r}
phyto.mat<-matrix(phyto$soilm, nrow=20)
```

```{r}
image.plot(phyto.mat, col=terrain.colors(24)) #original data
```



```{r}
phyto.polish<-medpolish(phyto.mat)
```


```{r}
phyto.polish$overall
```


```{r}
phyto.polish$row
```


```{r}
phyto.polish$col
```


```{r}
phyto.polish$residuals
```




```{r}
plot(phyto.polish$row)
```


```{r}
plot(phyto.polish$col)
```


```{r}
plot(phyto.polish$residuals)
```


```{r}
plot(phyto.polish$overall)
```



##Subtract the median polish residuals from the original data to capture the signal (trend) and plot it.

```{r}
phyto.trend<-phyto$soilm-phyto.polish$residuals
```

```{r}
image.plot(phyto.trend, col=terrain.colors(24)) #trend in the data
```



```{r}
image.plot(phyto.polish$residuals) #residuals
```



##Part 2: Constructing Correlograms

```{r}
dha=readOGR(dsn = "C:/Users/valer/Desktop/SIE512/Labs/Lab8/Lab82019",layer = "dhapol")
```


```{r}
dha.nbr <- poly2nb(dha) #Create a neighbors file with poly 2nb
```


```{r}
lung <- dha$LUNG_CAN_S
breast <- dha$BREAST_CN_
heartatt <- dha$MYC_INF_SM
```


##Create a correlogram for each of these variable

```{r}
mlung <- sp.correlogram(dha.nbr, lung, order=8, method="I", zero.policy = TRUE)
```

```{r}
mbreast <- sp.correlogram(dha.nbr, breast, order=8, method="I", zero.policy = TRUE)
```

```{r}
mheartatt <- sp.correlogram(dha.nbr, heartatt, order=8, method="I", zero.policy = TRUE)
```


##plot each correlogram

```{r}
plot(mlung, main="Moran's I Correlogram for Lung Cancer")
```


```{r}
plot(mbreast, main="Moran's I Correlogram for Breast Cancer")
```


```{r}
plot(mheartatt, main="Moran's I Correlogram for Heart Attack ")
```











