---
title: "Lab4"
author: "Valerie Kamgue"
date: "10/9/2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(spatstat)
library(splancs)
```

```{r}
uganda.ppp=dget("C:/Users/valer/Desktop/SIE512/Labs/Lab4/UgandaFile/ugandappp.txt")
```

```{r}
plot(uganda.ppp, pch=16)
```


```{r}
plot(gordon, pch=16)
```

```{r}
attributes (uganda.ppp) 
```

```{r}
attributes (gordon)
```

```{r}
np <- uganda.ppp$n
```

```{r}
np= gordon$n
```


```{r}
summary(uganda.ppp)
```

```{r}
summary(gordon)
```

```{r}
#lambda.uganda = summary(uganda.ppp)$intensity
#lambda.uganda
```

```{r}
#lambda.gordon = summary(gordon)$intensity
#lambda.gordon
```


```{r}
#nndist(uganda.ppp) 
```


```{r}
#nndist(gordon) 
```


###Function to compute the observed mean nearest neighbor distance

```{r}
mean.obsnnd <-function(onnd){
  #mean.value<- mean(onnd)
  value <- mean(onnd)
 # cat("The observed mean is",mean.value)
  return (value)
  }
```



```{r}
##mean.obsnnd(c(nndist(uganda.ppp)))
```

```{r}
###mean.obsnnd(c(nndist(gordon)))
```

###Function to compute the expected nearest neighbor mean distance for a CSR process
```{r}
mean.expnnd <- function(l) { 
	mean.value<- 1 / (2 * sqrt(l)) 
	#cat("The expected mean is",mean.value)
	return (mean.value)
	}
```

```{r}
##mean.expnnd(lambda.uganda)
```

```{r}
##mean.expnnd(lambda.gordon)
```


###Function to compute the variance and standard error for a CSR process

```{r}
Stderror <- function(l,m) { 
	varvalue<- (4-pi) / (4*l*pi*m)
	vstderror<- sqrt(varvalue)
	
	#cat("The variance is",varvalue, "\n") 
	
	#cat("The standard error is",sterror)
	return (vstderror)
	
	}
```

```{r}
##Stderror(lambda.uganda, uganda.ppp$n)
```

```{r}
##variance(lambda.gordon,gordon$n )
```

###Computation of zscore for Uganda and Gordon



```{r}
zscore <- function(ptpattern) {
  lambda<- summary(ptpattern)$intensity
  np <- ptpattern$n
  zvalue <- ((mean.obsnnd(c(nndist(ptpattern))) - mean.expnnd(lambda)))/Stderror(lambda, np)
return(zvalue)
}
```


```{r}
#zscore (uganda.ppp)
```

```{r}
#zscore (gordon)
```


##Function to compute P_value


```{r}
Pvalue <- function(valuezscore, alpha){
if (valuezscore < 0) { 
  pv <- pnorm(valuezscore)
  cat ("The P_value is", pv, "\n")
}else{
  pv <- 1-pnorm(valuezscore)
  cat ("The P_value is", pv, "\n")
}

if (pv< alpha){
  cat("Rejection of the CSR \n")
}else{
  cat("Failure to reject the CSR \n")}
}
```


```{r}
#Pvalue(zscore (uganda.ppp), 0.05)
```

```{r}
#Pvalue(zscore (gordon), 0.05)
```



```{r}
Main <- function (PointPattern, alpha) {
cat("The observed mean is", mean.obsnnd(c(nndist(PointPattern))),"\n") 
lambda <- summary(PointPattern)$intensity  
cat("The expected mean is", mean.expnnd(lambda),"\n")
Stde <- Stderror(lambda, PointPattern$n)
cat("The variance is", Stde*Stde,"\n" )
zval <- zscore (PointPattern)
cat("The Z-score is ", zval ,"\n")
Pvalue(zval, alpha)
} 
```

```{r}
Main(gordon, 0.05)
```
```{r}
Main(uganda.ppp, 0.05)
```




###Check the intensity of each point pattern

```{r}
initialbwuganda= bw.diggle(uganda.ppp)
```

```{r}
initialbwgordon= bw.diggle(gordon)
```

```{r}
plot(density(uganda.ppp, initialbwuganda))
```

```{r}
plot(density(gordon, initialbwgordon))
```

```{r}
otherbwuganda1= bw.CvL(uganda.ppp)
```


```{r}
otherbwuganda2= bw.scott(uganda.ppp)
```

```{r}
otherbwuganda3= bw.ppl(uganda.ppp)
```

```{r}
plot(density(uganda.ppp, otherbwuganda1))
```


```{r}
plot(density(uganda.ppp, otherbwuganda2))
```


```{r}
plot(density(uganda.ppp, otherbwuganda3))
```



```{r}
otherbwgordon1= bw.CvL(gordon)
```

```{r}
otherbwgordon2= bw.scott(gordon)
```

```{r}
otherbwgordon3= bw.ppl(gordon)
```

```{r}
plot(density(gordon, otherbwgordon1))
```

```{r}
plot(density(gordon, otherbwgordon2))
```

```{r}
plot(density(gordon, otherbwgordon3))
```


###Test second order effects using Ghat and Khat
##Run simulations for Ghat: 


```{r}
Gest(gordon, r=NULL, breaks=NULL, correction=c("rs", "km", "han"),domain=NULL)
```

```{r}
Ggordon <- Gest(gordon, r=NULL, breaks=NULL, correction=c("rs", "km", "han"),domain=NULL)
plot(Ggordon)
```


```{r}
Gest(uganda.ppp)
```


```{r}
Guganda <- Gest(uganda.ppp)
plot(Guganda)
```



```{r}
ugandaG.env<-envelope(uganda.ppp, fun=Gest, nsim=19, global =TRUE)
plot(ugandaG.env)
```


```{r}
gordonG.env<-envelope(gordon, fun=Gest, nsim=19, global =TRUE)
plot(gordonG.env)
```


###Run simulations for Lhat:

```{r}
Lgordon <- Lest(gordon)
plot(Lgordon)
```


```{r}
Luganda<- Lest(uganda.ppp)
plot(Luganda)
```


```{r}
ugandaL.env<-envelope(uganda.ppp, fun=Lest, nsim=19, global =TRUE)
plot(ugandaL.env)
```


```{r}
gordonL.env<-envelope(gordon, fun=Lest, nsim=19, global =TRUE)
plot(gordonL.env)
```


```{r}
plot(ugandaL.env, .-r ~ r)
```

