---
title: "Lab6"
author: "Valerie Kamgue"
date: "10/31/2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(maptools)
library(sp)
library(classInt)
library(RColorBrewer)
library(gstat)
library(lmtest)
library(tseries)
library(ggplot2)
library(rgdal)

```

```{r}
library(foreign)
```

####1.	Trend surface analysis

```{r}
calsoil <- readOGR(dsn="C:/Users/valer/Desktop/SIE512/Labs/Lab6/Lab6Data", layer="calcium")
```


##Plot the data using proportional symbols.

```{r}
plotvar <- calsoil$calcium
nclr=5
plotclr <- brewer.pal(nclr,"OrRd")
class <-classIntervals(plotvar, nclr, style ="equal",dataPrecision=1)
colcode <- findColours(class, plotclr)
max.symbol.size=4
min.symbol.size=.7
symbol.size <- ((plotvar-min(plotvar))/ (max(plotvar)-min(plotvar))*(max.symbol.size-min.symbol.size)+min.symbol.size)
plot(calsoil, pch=16, col=colcode, cex=symbol.size)
points(calsoil, cex=symbol.size)
title("Calcium Data")
legend(x = "bottomright", legend=names(attr(colcode, "table")),
    fill=attr(colcode, "palette"), cex=0.8, bty="n")

```




##Fit a linear trend surface model to the calcium data. 

```{r}
lm.ca<-lm(calcium ~ X+Y, data=calsoil)
summary(lm.ca)

```


```{r}
CX <- cor(calsoil$calcium, calsoil$X)
CX
```


```{r}
CY <- cor(calsoil$calcium, calsoil$Y)
CY
```


```{r}
Cdf<-as.data.frame(calsoil)
```


```{r}
ggplot(Cdf, aes(x=X, y=calcium)) + geom_point(col='red') +ylab("calcium")+ geom_smooth(method=lm, se=FALSE)
```



```{r}
ggplot(Cdf, aes(x=Y, y=calcium)) + geom_point(col='red') +ylab("calcium")+ geom_smooth(method=lm, se=FALSE)
```




```{r}
plot(lm.ca, which=1 ,pch=16)
```



```{r}
plot(lm.ca, which=2 ,pch=16)
```



```{r}
plot(lm.ca, which=5 ,pch=16)
```




```{r}
bptest(lm.ca)
jarque.bera.test(lm.ca$residuals)
shapiro.test(lm.ca$residuals)

```


```{r}
resids <-lm.ca$residual
calresids<-spCbind(calsoil, resids)

plotvar <- calresids$resids
nclr <- 5
plotclr <- brewer.pal(nclr,"RdBu") 
class <-classIntervals(plotvar, nclr, style ="sd")
colcode <- findColours(class, plotclr)
plot(calresids, pch=16, col=colcode, cex=1.5)
points(calresids, cex=1.5)
title("Calcium Residuals ")
legend(x = "bottomright", legend=names(attr(colcode, "table")),fill=attr(colcode, "palette"), cex=0.6, bty="n")

```


```{r}
anova(lm.ca)
```


##Fit a quadratic trend surface model

```{r}
calquad<-lm(calcium ~ X+Y+I(X^2)+I(Y^2)+I(X*Y), data=calsoil)
summary(calquad)

```



```{r}
anova(calquad)
```


##extract residual from quadratic model

```{r}
resids2 <-calquad$residual
calresids2<-spCbind(calsoil, resids2)

plotvar <- calresids2$resids2
nclr <- 5
plotclr <- brewer.pal(nclr,"RdBu") 
class <-classIntervals(plotvar, nclr, style ="sd")
colcode <- findColours(class, plotclr)
plot(calresids2, pch=16, col=colcode, cex=1.5)
points(calresids2, cex=1.5)
title("Calcium Residuals Quadratic Model ")
legend(x = "bottomright", legend=names(attr(colcode, "table")),fill=attr(colcode, "palette"), cex=0.6, bty="n")
```


```{r}
plot(calquad, which=1 ,pch=16)
```




```{r}
plot(calquad, which=2 ,pch=16)
```



```{r}
plot(calquad, which=5 ,pch=16)
```




##2. Examination of second order effects in Calcium Data
##Create a variogram of residuals from the liner trend model

```{r}
var1.calcium<-variogram(resids ~1, data=calresids)
```


```{r}
var1.calcium
```

```{r}
plot(var1.calcium,pch=16)
```



```{r}
plot(var1.calcium,pch=16,pl=T)
```







```{r}
covar.calcium<-variogram(resids ~1, data=calresids, covariogram=T)
plot(covar.calcium,pch=16)

```


##Run this function with your initial estimated values for sill, range, and nugget

```{r}
Ca1.vgm<-vgm(59 , "Sph", 219.7515 , 30)
```




```{r}
plot(var1.calcium, pch=16, Ca1.vgm)
```


##Fitting by Weighted Least Squares

```{r}
CaSph.fit<-fit.variogram(var1.calcium, Ca1.vgm)
CaSph.fit 
```


```{r}
plot(var1.calcium, pch=16, CaSph.fit)
```



##Next create an empirical variogram of residuals from the quadratic trend model, estimate variogram parameters for this model, fit the model using the weighted least squares fit.variogram function



```{r}
var2.calcium <- variogram(resids ~1, data=calresids2)
```


```{r}
var2.calcium
```


```{r}
plot(var2.calcium, pch=16)
```




```{r}
plot(var2.calcium,pch=16,pl=T)
```



```{r}
Ca2.vgm<-vgm(60 , "Sph", 230 , 30)
```

```{r}
plot(var2.calcium, pch=16, Ca2.vgm)
```


```{r}
CaSph2.fit<-fit.variogram(var2.calcium, Ca2.vgm)
CaSph2.fit
```

```{r}
plot(var2.calcium, pch=16, CaSph2.fit)
```





